\textit{Supporting Information for}: Geometric models reveal behavioral and neural signatures of transforming experiences into memories

### 2020-11-23 paper/supplementary_information.tex

\section*{Supplementary methods}
\subsection*{Optimizing topic model parameters}
In order to create accurate episode and recall models, we used an optimization method that was driven by our ability to explain hand-annotated memory performance metrics collected by \cite{ChenEtal17}.  In an earlier variant of our study~\citep{HeusMann18}, we used a grid search to compute the $\omega$ (episode sliding window duration, in scenes), $\rho$ (recall sliding window duration, in sentences), and $K$ (number of topics) that satisfied
\[
\argmax_{\omega, \rho, K} \left[\mathrm{corr}\left(\mathrm{corr}\left(\mu\left(\omega, \rho, K\right), \nu\left(\omega, \rho, K\right)\right), \theta\right)\right],
\]
where $\mathrm{corr}(\mu, \nu)$ is the per-participant correlation between the episode ($\mu$) and recall ($\nu$) topic proportions matrices, and $\theta$ is the per-participant hand-counted number of recalled scenes.  We searched over a grid of pre-specified values for each of these parameters; the resulting correlations are displayed in Figure~\ref{fig:paramsearch}.  The optimal parameters were $\omega = 50$, $\rho = 10$, and $K = 100$.  In our current paper we made a number of improvements to how we preprocessed text and fit topic models (see \textit{Methods}), but we carried the same optimal parameters forward from \cite{HeusMann18} without performing any additional optimization.

@@

The optimized model converged on 32 unique topics that were assigned non-zero weights.  We provide a list of the top ten highest-weighted words from each topic in Figure~\ref{fig:topics}.

@@

\subsection*{Feature importance analyses}
To determine the contribution of each feature to the temporal structure of the episode topic proportions matrix, we conducted a ``leave one out'' analysis.  Specifically, we compared the original episode's topic proportions matrix (created using all hand-annotated features from the 998 manually identified scenes spanning the \textit{Sherlock} episode; see \textit{Methods} for a full list of features) with topic proportions matrices created using all but one type of feature.  For each impoverished topic proportions matrix, we computed the timepoint-by-timepoint correlation matrix, and correlated the proximal diagonals from the upper triangle with those from the temporal correlation matrix of the feature-complete matrix (for details on how we isolated proximal temporal correlations, see \textit{Methods}).  Observing a lower correlation between an impoverished matrix (with a particular feature removed) and the feature-complete matrix would suggest that the held-out feature contributed more prominently to the full episode topic proportion matrix's temporal structure.  We found that hand-annotated narrative details played the greatest roll in determining the temporal structure of the episode, whereas the name of the character(s) in focus for each shot contributed least (Fig.~\ref{fig:feature-importance}A).

@@

Next, we sought to determine which annotated features contributed aspects of the episode's temporal structure that were preserved in participants' later recalls.  Specifically, we computed the timepoint-by-timepoint correlation matrix of the episode's topic proportions matrix, and correlated the proximal diagonals from its upper triangle with those from the timepoint-by-timepoint correlation matrices for each participant's recall topic proportions matrices (stretched via linear interpolation to have the same number of timepoints as the episode topic proportions matrix).  This yielded a single correlation coefficient for each participant.  We then repeated this analysis with each annotated feature held out in turn.  Observing a lower correlation between the episode and recall topic proportions matrices (constructed in the absence of a given feature) would indicate that participants utilize changes in that feature's content to discriminate between sections of the episode when organizing their recalls.  We found that hand-annotated narrative details were the most heavily utilized feature, whereas changes in the text present on-screen, the indoor/outdoor distinction between shots, the camera angle, the names of the various characters on screen, and the shot's location tended not to impact participants' recall structures (i.e., removing those features resulted in a \textit{greater} episode-recall correlation than including them; Fig.~\ref{fig:feature-importance}B).

We also wondered how the different types of features might relate.  For example, knowing which character is in focus during a given scene may also provide information about which character is speaking.  We computed topic proportions matrices from the annotations for each individual feature, in turn, and (using the same technique as in the above analyses) compared the proximal temporal correlation structure of each single-feature topic proportions matrix to each other, as well as to that of the full episode.  This provided additional confirmation that the full episode's temporal structure was largely driven by narrative details.  We also found that character-driven features (characters on screen, characters speaking, and characters in focus) were strongly correlated.  Other details, such as the presence or absence of music, led to very different topic proportions matrices (Fig.~\ref{fig:feature-importance}C).
%\FloatBarrier

@@

\begin{figure}[p!]
\centering
\includegraphics[width=\textwidth]{figs/matchmats}
\caption{\small \textbf{Episode-recall event correlation matrices.}  Each panel is in the same format as Figure~2G in the main text.  The yellow boxes mark the matched episode event for each recall event (i.e., the maximum correlation in each column).}
\label{fig:matchmats}
\end{figure}

>>>>

\section*{Supplementary methods}
\subsection*{Optimizing topic model parameters}
In order to create accurate episode and recall models, we used an optimization method that was driven by our ability to explain hand-annotated memory performance metrics collected by \cite{ChenEtal17}.  In an earlier variant of our study~\citep{HeusMann18}, we used a grid search to compute the $\omega$ (episode sliding window duration, in scenes), $\rho$ (recall sliding window duration, in sentences), and $K$ (number of topics) that satisfied
\[
\argmax_{\omega, \rho, K} \left[\mathrm{corr}\left(\mathrm{corr}\left(\mu\left(\omega, \rho, K\right), \nu\left(\omega, \rho, K\right)\right), \theta\right)\right],
\]
where $\mathrm{corr}(\mu, \nu)$ is the per-participant correlation between the episode ($\mu$) and recall ($\nu$) topic proportions matrices, and $\theta$ is the per-participant hand-counted number of recalled scenes.  We searched over a grid of pre-specified values for each of these parameters; the resulting correlations are displayed in Supplementary Figure~\ref{fig:paramsearch}.  The optimal parameters were $\omega = 50$, $\rho = 10$, and $K = 100$.  In our current paper we made a number of improvements to how we preprocessed text and fit topic models (see \textit{Methods}), but we carried the same optimal parameters forward from \cite{HeusMann18} without performing any additional optimization.

@@

The optimized model converged on 32 unique topics that were assigned non-zero weights.  We provide a list of the top ten highest-weighted words from each topic in Supplementary Figure~\ref{fig:topics}.

@@

\subsection*{Feature importance analyses}
To determine the contribution of each feature to the temporal structure of the episode topic proportions matrix, we conducted a ``leave one out'' analysis.  Specifically, we compared the original episode's topic proportions matrix (created using all hand-annotated features from the 998 manually identified scenes spanning the \textit{Sherlock} episode; see \textit{Methods} for a full list of features) with topic proportions matrices created using all but one type of feature.  For each impoverished topic proportions matrix, we computed the timepoint-by-timepoint correlation matrix, and correlated the proximal diagonals from the upper triangle with those from the temporal correlation matrix of the feature-complete matrix (for details on how we isolated proximal temporal correlations, see \textit{Methods}).  Observing a lower correlation between an impoverished matrix (with a particular feature removed) and the feature-complete matrix would suggest that the held-out feature contributed more prominently to the full episode topic proportion matrix's temporal structure.  We found that hand-annotated narrative details played the greatest roll in determining the temporal structure of the episode, whereas the name of the character(s) in focus for each shot contributed least (Supp. Fig.~\ref{fig:feature-importance}A).

@@

Next, we sought to determine which annotated features contributed aspects of the episode's temporal structure that were preserved in participants' later recalls.  Specifically, we computed the timepoint-by-timepoint correlation matrix of the episode's topic proportions matrix, and correlated the proximal diagonals from its upper triangle with those from the timepoint-by-timepoint correlation matrices for each participant's recall topic proportions matrices (stretched via linear interpolation to have the same number of timepoints as the episode topic proportions matrix).  This yielded a single correlation coefficient for each participant.  We then repeated this analysis with each annotated feature held out in turn.  Observing a lower correlation between the episode and recall topic proportions matrices (constructed in the absence of a given feature) would indicate that participants utilize changes in that feature's content to discriminate between sections of the episode when organizing their recalls.  We found that hand-annotated narrative details were the most heavily utilized feature, whereas changes in the text present on-screen, the indoor/outdoor distinction between shots, the camera angle, the names of the various characters on screen, and the shot's location tended not to impact participants' recall structures (i.e., removing those features resulted in a \textit{greater} episode-recall correlation than including them; Supp. Fig.~\ref{fig:feature-importance}B).

We also wondered how the different types of features might relate.  For example, knowing which character is in focus during a given scene may also provide information about which character is speaking.  We computed topic proportions matrices from the annotations for each individual feature, in turn, and (using the same technique as in the above analyses) compared the proximal temporal correlation structure of each single-feature topic proportions matrix to each other, as well as to that of the full episode.  This provided additional confirmation that the full episode's temporal structure was largely driven by narrative details.  We also found that character-driven features (characters on screen, characters speaking, and characters in focus) were strongly correlated.  Other details, such as the presence or absence of music, led to very different topic proportions matrices (Supp. Fig.~\ref{fig:feature-importance}C).
%\FloatBarrier

@@

\begin{figure}[p!]
\centering
\includegraphics[width=\textwidth]{figs/matchmats}
\caption{\small \textbf{Episode-recall event correlation matrices.}  Each panel is in the same format as Figure~\topicprops G in the main text.  The yellow boxes mark the matched episode event for each recall event (i.e., the maximum correlation in each column).}
\label{fig:matchmats}
\end{figure}
